Parity games in pushdown systems with parameters

### 2023-06-20 reachability/article.tex
upg
up

{\bf Games on pushdown systems} Pushdown systems can be used to model the behavior of recursive programs and,
as a result, a variety of 
 program analysis questions can be reduced to decision problems for games on pushdown systems.
They have applications for instance in inter-procedural control-flow analysis of recursive programs \cite{esparza1999automata, reps2005weighted}.

@@

{\bf Parametric pushdown automata or systems ?}
 Parametric pushdown systems provide a  formalism to
reason about 
recursive programs
making use of parametric constraints.
% Here, the stack can additionally be compared against parameters that can take unspecified word values. 
Recently, different variants of parametric pushdown systems have been introduced in the literature 
% [Hague’11]=[20], [Esparza, Ganty, Majumdar’13]=[15]
\cite{hague2011parameterised, esparza2016parameterized, fortin2017model} 
% In this work, following [21,20,15,27,12], we consider parametric pushdown systems
however as
parameterized asynchronous shared-memory systems.
These systems consist of a leader pushdown automaton
and arbitrarily many identical contributor pushdown systems, communicating via a shared memory
in the form of a register which can take finitely many values.
This variant have been shown to have applications
to the dataflow analysis of concurrent programs \cite{kahlon2008parameterization}. 
We consider here a different approach to extending pushdown systems with parameters, by allowing them to test equality of the stack content against parameters.
% Pushdown systems which can test the stack content against regular-expressions are equivalent to pushdown systems, so pushdown systems which can test the stack content against specific words are too.
Perhaps a similar model consists in 
pushdown systems 
with transitions that are conditioned by regular conditions on their stack content, which
can in particular
test the
stack content against
specific words. 
They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages).
They have \textcolor{red}{ phrases trop longues}
been used to establish that CTL$^*$ model checking remains decidable
when the formulas are allowed to include regular predicates on the stack content,
and to obtain 
model checking algorithms for LTL and CTL$^*$ model checking
for
pushdown systems \cite{finkel1997direct}.
Pushdown systems 
%with regularity condition on the stack
with transitions that are conditioned by regular conditions on their stack content
can be viewed as
a special case of stack automata as seen in \cite{hopcroft1969formal}.
Instead
of checking the stack content against regular conditions, or, in a more limited manner,
against specified word values,
we consider
checks against unspecified word values that can be instantiated using parameters.
This allows for example to ensure equality between two stack contents appearing in two distincts configurations in a run.

@@

\subsubsection*{Overview}

In Section 2, we provide general notations and preliminary definitions. 
Section 3 will deal with the introduction of  parametric pushdown games.
% Section 4 is devoted to using pebble alternating tree systems with parity acceptance condition to solve parametric pushdown system parity games. We leave the proof of decidability itself  for Section 5.
Section 4 is Reduction to higher-order pushdown systems
Section 5 at last shows a nonelementary lower bound for the problem.

@@

A partial play $\pi = s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_n$ is {\em consistent} with a strategy $\sigma_i$ if sequences of configurations that end in $S_i$ along this play have all successors according to the strategy, i.e.
if for all $j\in[0,n-1]$, for all element $v_j \in V_i$ in the sequence $\pi$,
$v_{j+1} = \sigma_i(v_j)$.
 Given a configuration $s$,   a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1$ for player $1$, the play starting with $s$ that is consistent with both strategies is unique and is called the {\em resulting play} of $\sigma_0$ and $\sigma_1$ starting from $s$. It is 
denoted by $\pi(s, \sigma_0, \sigma_1)$.\\

@@

	For the decidability upper bound, we

>>>>

{\bf Games on pushdown systems} Pushdown systems can be used to model the behavior of recursive programs.
As a result, a variety of 
 program analysis questions can be reduced to decision problems for games on pushdown systems.
They have applications for instance in inter-procedural control-flow analysis of recursive programs \cite{esparza1999automata, reps2005weighted}.

@@

{\bf Parametric pushdown systems}
 Parametric pushdown systems provide a  formalism to
reason about 
recursive programs
making use of parametric constraints.
% Here, the stack can additionally be compared against parameters that can take unspecified word values. 
Recently, different variants of parametric pushdown systems have been introduced in the literature 
% [Hague’11]=[20], [Esparza, Ganty, Majumdar’13]=[15]
\cite{hague2011parameterised, esparza2016parameterized, fortin2017model} 
% In this work, following [21,20,15,27,12], we consider parametric pushdown systems
however as
parameterized asynchronous shared-memory systems.
These systems consist of a leader pushdown automaton
and arbitrarily many identical contributor pushdown systems, communicating via a shared memory
in the form of a register which can take finitely many values.
This variant have been shown to have applications
to the dataflow analysis of concurrent programs \cite{kahlon2008parameterization}. 
We consider here a different approach to extending pushdown systems with parameters, by allowing them to test equality of the stack content against parameters.
% Pushdown systems which can test the stack content against regular-expressions are equivalent to pushdown systems, so pushdown systems which can test the stack content against specific words are too.
Perhaps a similar model consists in 
pushdown systems 
with transitions that are conditioned by regular conditions on their stack content, which
can in particular
test the
stack content against
specific words. 
They can be used to ensure that some word over the stack alphabet appears in the configurations of a run, but only for a specific value (or, rather, specific regular languages).
They have \textcolor{red}{ phrases trop longues}
been used to establish that CTL$^*$ model checking remains decidable
when the formulas are allowed to include regular predicates on the stack content,
and to obtain 
model checking algorithms for LTL and CTL$^*$ model checking
for
pushdown systems \cite{finkel1997direct}.
Pushdown systems 
%with regularity condition on the stack
with transitions that are conditioned by regular conditions on their stack content
can be viewed as
a special case of stack automata as seen in \cite{hopcroft1969formal}.
Instead
of checking the stack content against regular conditions, or, in a more limited manner,
against specified word values,
we consider
checks against unspecified word values that can be instantiated using parameters.
This allows for example to ensure equality between two stack contents appearing in two distincts configurations in a run.

@@

{\bf Overview. }
In Section 2, we provide general notations and preliminary definitions. 
Section 3 will deal with the introduction of  parametric pushdown games.
% Section 4 is devoted to using pebble alternating tree systems with parity acceptance condition to solve parametric pushdown system parity games. We leave the proof of decidability itself  for Section 5.
Section 4 is Reduction to higher-order pushdown systems
Section 5 at last shows a nonelementary lower bound for the problem.

@@

A partial play $\pi = s_0 \rightarrow s_1 \rightarrow \ldots \rightarrow s_n$ is {\em consistent} with a strategy $\sigma_i$ if sequences of configurations that end in $S_i$ along this play have all successors according to the strategy, i.e.
if for all $j\in[0,n-1]$, for all element $v_j \in V_i$ in the sequence $\pi$,
$v_{j+1} = \sigma_i(v_j)$.
 Given a configuration $s$,   a strategy $\sigma_0$ for player $0$ and a strategy $\sigma_1$ for player $1$, the play starting with $s$ that is consistent with both strategies is unique and is called the {\em resulting play} of $\sigma_0$ and $\sigma_1$ starting from $s$. It is 
denoted by $\pi(s, \sigma_0, \sigma_1)$.

@@

	For the upper bound, we replaced parameters by pebbles acting as registers, leading to a more general model, and then used a reducton to higher-order pushdown systems parity games to solve parity games. Since solving parity games on higher-order pushdown systems with level $n$ stack is $n$-$\EXP$-complete \cite{ Cach03, cachat2007complexity}, this provides an $(n+1)$-$\EXP$ upper bound for solving parametric parity games on parametric pushdown systems with $n$ parameters.
